---
title: "Email Report"
author: "Justin Wininger"
date: "`r Sys.Date()`"
output: word_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, warning = FALSE, message = F}
knitr::opts_chunk$set(echo = FALSE, fig.width = 8, fig.height = 5, dpi = 300)
remove(list = ls())
library(readxl)
library(tidyverse)
library(lubridate)
library(hrbrthemes)
library(viridis)
library(mosaic)
library(factoextra)

emails <- read_excel("PGI master 2.xlsx")


emails <- emails %>%
  mutate(`Open Rate` = `Open Rate` * 100)

emails <- emails %>%
  mutate(`Click Rate` = `Click Rate` * 100)

emails <- emails %>%
  mutate(CTOR = CTOR * 100)
```

```{r message = F, warning = F}
emails$Day <- as.Date(emails$`Sent Date`)
emails$Day <- weekdays(emails$Day)
day_order <- c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")
emails$Day <- factor(emails$Day, levels = day_order, ordered = TRUE)


emails <- emails %>% filter(`Total Recipients` > 10)

emails$Covid <- ifelse(emails$`Sent Date` < as.Date("2020-03-01"), "Pre-COVID", "Post-COVID")

covid_order <- c("Pre-COVID", "Post-COVID")
emails$Covid <- factor(emails$Covid, levels = covid_order, ordered = TRUE)

counts <- emails %>% 
  group_by(Day) %>%
  tally()

average_recipients <- emails %>%
  group_by(Client) %>%
  summarise(recipientList = median(`Total Recipients`))

average_recipients$Size <- ifelse(average_recipients$recipientList >= 2500, "Large", "Small")

average_recipients <- subset( average_recipients, select = -recipientList )

emails <- left_join(emails, average_recipients, by = 'Client')

large <- emails %>% filter(Size == 'Large')
small <- emails %>% filter(Size == 'Small')

countsL <- large %>% 
    group_by(Client) %>%
    summarize(Count = n())

countsS <- small %>% 
    group_by(Client) %>%
    summarize(Count = n())

countsSL <- full_join(countsS, countsL, by = 'Client')

emailsMF <- emails %>% filter(Day != 'Sunday')

average_recipients <- emails %>%
  group_by(Client) %>%
  summarise(recipientList = median(`Total Recipients`))

average_recipients$Size <- ifelse(average_recipients$recipientList >= 2500, "Large", "Small")

size <- emails %>%
  group_by(Size)%>%
  summarise(
    openRate=median(`Open Rate`),
    CTOR=median(CTOR)
    )

size <- subset( size, select = -Size )

chisqS <- chisq.test(size)
#p = 0.3323

emailsBC <- subset(emailsMF, Covid == "Pre-covid")
emailsAC <- subset(emailsMF, Covid == "Post-covid")

emails <- emails %>% mutate(`Subject Length` = nchar(Subject))

SL <- lm(`Open Rate` ~ `Subject Length`, data = emails)

sSL <- summary(SL)

sSLo <- sSL[["coefficients"]]

has_punctuation <- function(text) {
  grepl("[[:punct:]]", text)
}

# Add a new column 'contains_punctuation' based on the presence of punctuation
emails <- emails %>%
  mutate(Punctuation = ifelse(has_punctuation(Subject), "Yes", "No"))

counts2 <- emails %>% 
  group_by(Punctuation) %>%
  tally()

# Function to check for specific punctuation marks
has_period <- function(text) {
  grepl("\\.", text)
}

has_colon <- function(text) {
  grepl(":", text)
}

has_exclamation <- function(text) {
  grepl("!", text)
}

has_question <- function(text) {
  grepl("\\?", text)
}

has_comma <- function(text) {
  grepl("\\,", text)
}

has_hyphen <- function(text) {
  grepl("\\-", text)
}

has_ampersand <- function(text) {
  grepl("\\&", text)
}

has_quote <- function(text) {
  grepl('\\"', text)
}

has_em <- function(text) {
  grepl('\\—', text)
}

has_em2 <- function(text) {
  grepl('\\–', text)
}

has_slash <- function(text) {
  grepl('\\/', text)
}

# Add new columns for specific punctuation marks
emails <- emails %>%
  mutate(
    Period = ifelse(has_period(Subject), "Yes", "No"),
    Colon = ifelse(has_colon(Subject), "Yes", "No"),
    ExclamationPoint = ifelse(has_exclamation(Subject), "Yes", "No"),
    QuestionMark = ifelse(has_question(Subject), "Yes", "No"),
    Comma = ifelse(has_comma(Subject), "Yes", "No"),
    Hyphen = ifelse(has_hyphen(Subject), "Yes", "No"),
    Ampersand = ifelse(has_ampersand(Subject), "Yes", "No"),
    Quote = ifelse(has_quote(Subject), "Yes", "No"),
    Em = ifelse(has_em(Subject), "Yes", "No"),
    Em2 = ifelse(has_em2(Subject), "Yes", "No"),
    Slash = ifelse(has_slash(Subject), "Yes", "No")
  )

haveP <- emails %>%
  filter(Period == "Yes") %>%
  mutate(Punc = "Period")
haveC <- emails %>%
  filter(Colon == "Yes") %>%
  mutate(Punc = "Colon")
haveE <- emails %>%
  filter(ExclamationPoint == "Yes")%>%
  mutate(Punc = "Exclamation\nPoint")
haveQ <- emails %>%
  filter(QuestionMark == "Yes")%>%
  mutate(Punc = "Question\nMark")
haveN <- emails %>%
  filter(Punctuation == "No")%>%
  mutate(Punc = "None")
haveCom <- emails %>%
  filter(Comma == 'Yes')%>%
  mutate(Punc = 'Comma')
haveD <- emails %>%
  filter(Hyphen == 'Yes' | Em == 'Yes' | Em2 == 'Yes')%>%
  mutate(Punc = 'Dash')
haveAmp <- emails %>%
  filter(Ampersand == 'Yes')%>%
  mutate(Punc = 'Ampersand')
haveQuo <- emails %>%
  filter(Quote == 'Yes')%>%
  mutate(Punc = 'Quotes')
haveSlash <- emails %>%
  filter(Slash == 'Yes')%>%
  mutate(Punc = 'Slash')


tmp <- emails %>% filter(Punctuation == 'Yes' & (Hyphen == 'Yes' | Em == 'Yes' | Em2 == 'Yes'))

havePunc <- do.call("rbind", list(haveP, haveC, haveE, haveQ, haveN, haveCom, haveD, haveAmp, haveQuo, haveSlash))

havePunc <- havePunc %>% 
  group_by(Punc) %>%
  filter(n() >= 15)

analyze_last_char <- function(text) {
  last_char <- substr(text, nchar(text), nchar(text))
  if (last_char == ".") {
    return("Period")
  } else if (last_char == "!") {
    return("Exclamation")
  } else if (last_char == "?") {
    return("Question")
  } else if (last_char == ":") {
    return("Colon")
  } else if (last_char == "&") {
    return("Ampersand")
  } else if (last_char == ",") {
    return("Comma")
  } else if (last_char == "/") {
    return("Slash")
  } else if (last_char == '"') {
    return("Quotes")
  } else if (last_char == "-" | last_char == "—" | last_char == "–") {
    return("Dash")
  } else {
    return("No Punctuation")
  }
}

# Create new column with analysis of last character
emails$lastChar <- sapply(emails$Subject, analyze_last_char)

emailsLC <- emails %>% 
  filter(lastChar == 'Period' | lastChar == 'No Punctuation') 

counts6 <- emails %>% 
  group_by(lastChar) %>%
  tally()

noPunc <- emails %>% filter(Punctuation == "No")
noPunc <- noPunc %>%
  mutate (Source = "No Punctuation")
emailsNP <- emailsLC %>% 
  filter(lastChar == 'No Punctuation') %>%
  mutate (Source = "No Ending Punctuation")
emailsP <- emailsLC %>% 
  filter(lastChar == 'Period')%>%
  mutate (Source = "Ending Period") 
tmp <- rbind(emailsNP, emailsP)
both <- rbind(tmp, noPunc)

countsBoth <- both %>% 
  group_by(Source) %>%
  tally()

both2 <- rbind(emailsP, noPunc)

countsBoth2 <- both2 %>% 
  group_by(Source) %>%
  tally()

countsSize <- emails %>% 
  group_by(Size) %>%
  tally()

emails <- emails %>%
  group_by(Client, Subject) %>%
  mutate(Duplicate = if_else(n_distinct(`Sent Date`) > 1, "Yes", "No")) %>%
  ungroup()

countsDup <- emails %>% 
  group_by(Duplicate) %>%
  tally()

analyze_keywords <- function(data, subject_col, keywords) {
  emails <- emails %>%
    mutate(across(all_of(subject_col), tolower, .names = "lower_{col}"))
  
  for (keyword_group in keywords) {
    # Split the keywords if there are synonyms
    keyword_synonyms <- unlist(strsplit(keyword_group, "/"))
    base_keyword <- keyword_synonyms[1]
    
    # Create a new column name
    new_col_name <- paste0("have", base_keyword)
    
    # Check if any of the synonyms are in the subject column
    emails[[new_col_name]] <- ifelse(grepl(paste(keyword_synonyms, collapse = "|"), emails[[paste0("lower_", subject_col)]]), "Yes", "No")
  }
  
  # Check for numbers 1-10 in both numeric and written form
  numbers_written <- c("one", "two", "three", "four", "five", "six", "seven", "eight", "nine", "ten")
  numbers_numeric <- as.character(1:10)
  all_numbers <- c(numbers_written, numbers_numeric)
  emails[["haveNumber"]] <- ifelse(grepl(paste(all_numbers, collapse = "|"), emails[[paste0("lower_", subject_col)]]), "Yes", "No")
  
  # Remove the temporary lowercase subject column
  emails <- emails %>%
    select(-starts_with("lower_"))
  
  return(emails)
}

# Example usage
# Assuming `df` is your dataset and `Subject` is the column to analyze
keywords <- c("gift/gifts", "legacy/legacies", "easy/smart/simple", "you/your/yours", 
              "impact/effect/affect", "option/options", "meaningful", "wise", 
              "support", "year-end", "give", "reward/rewarding", "will", 
              "support", "tax/taxes", "us/we", "help", "plan/planning")

# Applying the function to the dataset
emails <- analyze_keywords(emails, "Subject", keywords)

haveGift <- emails %>%
  filter(havegift == "Yes") %>%
  mutate(Keyword = "Gift(s)")
haveLegacy <- emails %>%
  filter(havelegacy == "Yes") %>%
  mutate(Keyword = "Legacy(ies)")
haveEasy <- emails %>%
  filter(haveeasy == "Yes")%>%
  mutate(Keyword = "Easy / Smart / Simple")
haveYou <- emails %>%
  filter(haveyou == "Yes")%>%
  mutate(Keyword = "You / Your / Yours")
haveImpact <- emails %>%
  filter(haveimpact == "Yes")%>%
  mutate(Keyword = "Impact / Effect")
haveOption <- emails %>%
  filter(haveoption == 'Yes')%>%
  mutate(Keyword = 'Option(s)')
haveMeaningful <- emails %>%
  filter(havemeaningful == 'Yes')%>%
  mutate(Keyword = 'Meaningful')
haveWise <- emails %>%
  filter(havewise == 'Yes')%>%
  mutate(Keyword = 'Wise')
haveSupport <- emails %>%
  filter(havesupport == 'Yes')%>%
  mutate(Keyword = 'Support')
haveYearEnd <- emails %>%
  filter(`haveyear-end` == 'Yes')%>%
  mutate(Keyword = 'Year-End')
haveNumber <- emails %>%
  filter(haveNumber == "Yes") %>%
  mutate(Keyword = "A Number")
haveGive <- emails %>%
  filter(havegive == "Yes") %>%
  mutate(Keyword = "Give")
haveReward <- emails %>%
  filter(havereward == "Yes")%>%
  mutate(Keyword = "Reward(ing)")
haveWill <- emails %>%
  filter(havewill == "Yes")%>%
  mutate(Keyword = "Will")
haveTax <- emails %>%
  filter(havetax == "Yes")%>%
  mutate(Keyword = "Tax(es)")
haveUs <- emails %>%
  filter(haveus == 'Yes')%>%
  mutate(Keyword = 'Us / We')
haveHelp <- emails %>%
  filter(havehelp == 'Yes')%>%
  mutate(Keyword = 'Help')
havePlan <- emails %>%
  filter(haveplan == 'Yes')%>%
  mutate(Keyword = 'Plan(ning)')

keywords <- do.call("rbind", list(haveGift, haveLegacy, haveEasy, haveYou, haveImpact, haveOption, haveMeaningful, haveWise, haveSupport, haveYearEnd, haveNumber, haveGive, haveReward, haveWill, haveTax, haveUs, haveHelp, havePlan))

# emails <- emails %>% 
#   mutate(
#     `Product Type` = fct_recode(as.factor(`Product Type`),
#                       `Donor-Centric\nNewsletter` = "Donor-Centric Newsletter",
#                       `Donor-Centric\nPostcard` = "Donor-Centric Postcard",
#                       `Donor-Centric\nTarget Mailer` = "Donor-Centric Target Mailer",
#                       `Professional\nAdvisor\nNewsletter` = "Professional Advisor Newsletter")
#   )
# 
# product_order <- c("Donor-Centric\nPostcard", "Donor-Centric\nTarget Mailer", "Visions", "Donor-Centric\nNewsletter", "Professional\nAdvisor\nNewsletter")
# emails$`Product Type` <- factor(emails$`Product Type`, levels = product_order, ordered = TRUE)
# 
# techniques <- emails %>% filter(`Product Type` == 'Professional\nAdvisor\nNewsletter')
# 
# countsTechniques <- techniques %>% 
#   group_by(Day) %>%
#   tally()
# 
# emails <- emails %>%
#   group_by(Client, Subject) %>%
#   mutate(Duplicate = if_else(n_distinct(`Sent Date`) > 1, "Yes", "No")) %>%
#   ungroup()
# 
# emails$Techniques <- ifelse(emails$`Product Type` == "Professional\nAdvisor\nNewsletter", "Yes", "No")
```

# Introduction

This is a report of 365 marketing email campaigns containing almost 290,000 individual emails sent by PGI Partners, Inc. and Endowment Development Services, Inc. since 2008. This report will look at several statistics of each campaign, including open rate and click-to-open rate. This report will almost exclusively use the median instead of the mean. The median, being resistant to outliers, is a much better representation of the "typical" case.

# Plots

Below are several plots that display different relationships between click-to-open rate and open rate, as well as subject lines and day of the week. Most of these plots will be box-and-whisker plots, where every line is significant. Any dot is an outlier, and the smaller lines at the very bottom and top represent the minimum and maximum, respectively, excluding outliers. The lines at the bottom and top of the box represent the first and third quartiles (25th and 75th percentiles, respectively), the lime colored line represents the median, and the white text is the actual median. The number above each box is the sample size. In general, results with a sample size lower than 10 were excluded, but sometimes they were not omitted. 

```{r message=FALSE, warning=FALSE}
countsMF <- emailsMF %>% 
  group_by(Day) %>%
  tally()

ggplot(data=emailsMF, mapping=aes(x=Day, y=`Open Rate`, fill = Day))+
  stat_boxplot(geom ='errorbar', width = .25) + 
  geom_boxplot(fill = c('#003a5d', '#0091b3', '#003a5d', '#0091b3', '#003a5d'))+
  theme_ipsum(axis_title_size = 12) +
    theme(
      legend.position="none",
    )+
  labs(title = 'Open Rate by Day of the Week')+
  stat_summary(fun = median, geom = "crossbar", width = .75, color = "#a9ad00", size = .4)+
  stat_summary(fun = median, geom = "text", aes(label = paste0(round(..y.., 2), "%")), 
               position = position_dodge(width = 0.75), vjust = -0.5, color = "white")+
  geom_text(data = counts,
            aes(Day, Inf, label = n), vjust = 1)

ggplot(data=emailsMF, mapping=aes(x=Day, y=CTOR, fill = Day))+
  stat_boxplot(geom ='errorbar', width = .25) +
  geom_boxplot(fill = c('#003a5d', '#0091b3', '#003a5d', '#0091b3', '#003a5d'))+
  theme_ipsum(axis_title_size = 12) +
    theme(
      legend.position="none",
    )+
  labs(title = 'Click-to-Open Rate by Day of the Week')+
  stat_summary(fun = median, geom = "crossbar", width = .75, color = "#a9ad00", size = .4)+
  stat_summary(fun = median, geom = "text", aes(label = paste0(round(..y.., 2), "%")), 
               position = position_dodge(width = 0.75), vjust = -0.5, color = "white")+
  geom_text(data = countsMF,
            aes(Day, Inf, label = n), vjust = 1)

days <- emails %>% 
  group_by(Day) %>% 
  summarise(
    openRate=round(median(`Open Rate`), digits =4),
    CTOR=round(median(CTOR), digits =4)
    )
```

```{r warning = F, message = F}
countsMF2 <- emailsMF %>% 
  group_by(Day, Covid) %>%
  tally()

emailsMF %>%
  ggplot(aes(Day, `Open Rate`, fill = Day,))+
  stat_boxplot(geom ='errorbar', width = .25) +
  geom_boxplot(fill = c('#003a5d', '#0091b3', '#003a5d', '#0091b3', '#003a5d', '#003a5d', '#0091b3', '#003a5d', '#0091b3', '#003a5d'))+
  theme_ipsum(axis_title_size = 12) +
    theme(
      legend.position="none",
      axis.text.x = element_text(angle = 90, hjust=1)
    )+
  labs(title = 'Open Rate by Day of the Week Relative to COVID-19')+
  facet_wrap( ~ Covid)+
  stat_summary(fun = median, geom = "crossbar", width = .75, color = "#a9ad00", size = .4)+
  stat_summary(fun = median, geom = "text", aes(label = paste0(round(..y.., 1), "%")),
               position = position_dodge(width = 0.75), vjust = -0.5, color = "white")+
  geom_text(data = countsMF2,
            aes(Day, Inf, label = n), vjust = 1)

daysC <- emails %>% 
  group_by(Day, Covid) %>% 
  summarise(
    CTOR=round(median(CTOR), digits =4),
    )

daysC <- daysC %>%
  pivot_wider(id_cols = Covid, names_from = Day, values_from = CTOR)
```

```{r}
emailsMF %>%
  ggplot(aes(Day, CTOR, fill = Day,))+
  stat_boxplot(geom ='errorbar', width = .25) +
  geom_boxplot(fill = c('#003a5d', '#0091b3', '#003a5d', '#0091b3', '#003a5d', '#003a5d', '#0091b3', '#003a5d', '#0091b3', '#003a5d'))+
  theme_ipsum(axis_title_size = 12) +
    theme(
      legend.position="none",
      axis.text.x = element_text(angle = 90, hjust=1)
    )+
  labs(title = 'Click-to-Open Rate by Day of the Week Relative to COVID-19',
       y= 'Click-to-Open Rate')+
  facet_wrap( ~ Covid)+
  stat_summary(fun = median, geom = "crossbar", width = .75, color = "#a9ad00", size = .4)+
  stat_summary(fun = median, geom = "text", aes(label = paste0(round(..y.., 1), "%")),
               position = position_dodge(width = 0.75), vjust = -0.5, color = "white")+
  geom_text(data = countsMF2,
            aes(Day, Inf, label = n), vjust = 1)
```

Friday seems to be far and away the best day to send emails, full stop. Monday is a good second option, but Friday is the leader in open and click-to-open rates by a large margin.

```{r warning = F, message = F}
# set.seed(123)
# 
# emails2 <- emails %>%
#   select(`Subject Length`, `Open Rate`)
# 
# fviz_nbclust(emails2, kmeans)
# 
# kmeans_result <- kmeans(emails2, centers = 3)
# 
# # Add cluster assignment to the data
# emails2 <- emails2 %>% mutate(cluster = as.factor(kmeans_result$cluster))
# 
# # Plot the scatterplot with clusters
# ggplot(emails2, aes(x = `Subject Length`, y = `Open Rate`, color = cluster)) +
#   geom_point(size = 3) +
#   labs(title = "K-means Clustering",
#        x = "X-axis",
#        y = "Y-axis") +
#   theme_minimal()+
#   geom_smooth(method=lm, se =F)

ggplot(data=emails, mapping=aes(x=`Subject Length`, y=`Open Rate`))+
  geom_point(color = '#0091b3')+
  theme_ipsum(axis_title_size = 12) +
    theme(
      legend.position="none",
    )+
  geom_smooth(se=FALSE, color = '#003a5d', formula = y~x)+
  labs(title = 'Open Rate by Subject Length',
       x = 'Subject Length (characters)')

ggplot(data=emails, mapping=aes(x=`Subject Length`, y=CTOR))+
  geom_point(color = '#0091b3')+
  theme_ipsum(axis_title_size = 12) +
    theme(
      legend.position="none",
    )+
  geom_smooth(se=FALSE, color = '#003a5d', formula = y~x)+
  labs(title = 'Click-to-Open Rate by Subject Length',
       x = 'Subject Length (characters)')

ORSL <- lm(`Open Rate` ~ `Subject Length`, emails)
sORSL <- summary(ORSL)
mORSL <- sORSL[["coefficients"]]

CTORSL <- lm(`Open Rate` ~ CTOR, emails)
sCTORSL <- summary(ORSL)
mCTORSL <- sORSL[["coefficients"]]

ORrf <- mplot(CTORSL, which=1)
ORqq <- mplot(CTORSL, which=2)

CTORrf <- mplot(CTORSL, which=1)
CTORqq <- mplot(CTORSL, which=2)
```

In general, as subject length increases, so do open and click-to-open rates. Both those relationships are statistically significant. It is important to stay within the bounds of the data, however. A ten thousand character subject would not go over that well. The data seem to show that subject lines between 40 and 70 characters perform the best for both open and click-to-open rates.

```{r warning = F, message = F}
punctuation <- emails %>%
  group_by(Punctuation)%>%
  summarise(
    openRate=round(median(`Open Rate`), digits =2),
    CTOR=round(median(CTOR), digits =2)
    )

punctuation <- punctuation[,-1]

chisqP <- chisq.test(punctuation)
#p = 0.7703
```

```{r warning = F, message = F}
counts7 <- havePunc %>% 
  group_by(Punc) %>%
  tally()

havePunc %>%
  ggplot(aes(x = reorder(Punc, `Open Rate`, FUN = median), y = `Open Rate`, fill = Punc))+
  stat_boxplot(geom ='errorbar', width = .25) +
  geom_boxplot(fill = c('#003a5d', '#0091b3', '#003a5d', '#0091b3', '#003a5d'))+
  theme_ipsum(axis_title_size = 12) +
  theme(
    legend.position="none",
    )+
  labs(title = "Open Rate by Type of Punctuation",
       x = 'Punctuation')+
  stat_summary(fun = median, geom = "crossbar", width = .75, color = "#a9ad00", size = .4)+
  stat_summary(fun = median, geom = "text", aes(label = paste0(round(..y.., 2), "%")), 
               position = position_dodge(width = 0.75), vjust = -0.5, color = "white")+
  geom_text(data = counts7,
            aes(Punc, Inf, label = n), vjust = 1)

havePunc %>%
  ggplot(aes(x = reorder(Punc, CTOR, FUN = median), y = CTOR, fill = Punc))+
  stat_boxplot(geom ='errorbar', width = .25) +
  geom_boxplot(fill = c('#003a5d', '#0091b3', '#003a5d', '#0091b3', '#003a5d'))+
  theme_ipsum(axis_title_size = 12) +
  theme(
    legend.position="none",
    )+
  labs(title = "Click-to-Open Rate by Type of Punctuation",
       x = 'Punctuation')+
  stat_summary(fun = median, geom = "crossbar", width = .75, color = "#a9ad00", size = .4)+
  stat_summary(fun = median, geom = "text", aes(label = paste0(round(..y.., 2), "%")), 
               position = position_dodge(width = 0.75), vjust = -0.5, color = "white")+
  geom_text(data = counts7,
            aes(Punc, Inf, label = n), vjust = 1)

havePuncTable <- havePunc %>%
  group_by(Punc) %>%
  summarize(
    CTOR = round(median(CTOR), digits =2)
  )
```

The plots on punctuation double dip on data. For example, a subject line with a colon and period would be counted twice, once in each box. So, just because period has the highest median open and click-to-open rates does not mean one should exclude every other type of punctuation.

```{r message = F, warning = F}
counts3 <- emails %>% 
  group_by(lastChar) %>%
  tally()

emails %>%
  ggplot(aes(x = reorder(lastChar, `Open Rate`, FUN = median), y = `Open Rate`, fill = lastChar)) +
  stat_boxplot(geom ='errorbar', width = .25) +
  geom_boxplot(fill = c('#003a5d', '#0091b3', '#003a5d', '#0091b3')) +
    scale_fill_viridis(discrete = TRUE) +
    theme_ipsum(axis_title_size = 12) +
    theme(
      legend.position="none",
    )+
  labs(x= 'Last Character',
       y= 'Open Rate',
       title = ' Open Rate by Last Character')+
  scale_fill_manual(values=c('#003a5d', '#0091b3','#003a5d', '#0091b3'))+
  stat_summary(fun = median, geom = "crossbar", width = .75, color = "#a9ad00", size = .4)+
  stat_summary(fun = median, geom = "text", aes(label = paste0(round(..y.., 2), "%")), 
               position = position_dodge(width = 0.75), vjust = -0.5, color = "white")+
  geom_text(data = counts3,
            aes(lastChar, Inf, label = n), vjust = 1)

emails %>%
  ggplot(aes(x = reorder(lastChar, CTOR, FUN = median), y = CTOR, fill = lastChar)) +
  stat_boxplot(geom ='errorbar', width = .25) +
  geom_boxplot(fill = c('#003a5d', '#0091b3', '#003a5d', '#0091b3')) +
    scale_fill_viridis(discrete = TRUE) +
    theme_ipsum(axis_title_size = 12) +
    theme(
      legend.position="none",
    )+
  labs(x= 'Last Character',
       y= 'Click-to-Open Rate',
       title = ' Click-to-Open Rate by Last Character')+
  # scale_fill_manual(values=c('#003a5d', '#0091b3'))+
  stat_summary(fun = median, geom = "crossbar", width = .75, color = "#a9ad00", size = .4)+
  stat_summary(fun = median, geom = "text", aes(label = paste0(round(..y.., 2), "%")), 
               position = position_dodge(width = 0.75), vjust = -0.5, color = "white")+
  geom_text(data = counts3,
            aes(lastChar, Inf, label = n), vjust = 1)
  
modelLC <- lm(`Open Rate` ~ lastChar, emails)
anovaLC <- anova(modelLC)
#p = 1.268868e-14

LCTable <- emailsLC %>%
  group_by(lastChar)%>%
  summarise(
    COTR = round(median(CTOR), digits =2)
  )
```

Looking at the last character in the subject line, exclamation points appear to do really well. They lead in both open and click-to-open rate.

```{r warning = F, message = F}
dupes <- emails %>% 
  group_by(Duplicate) %>% 
  summarise(
    openRate=round(median(`Open Rate`), digits =4),
    CTOR=round(median(CTOR), digits =4)
    )

dupes <- dupes[,-1]

chisqD <- chisq.test(dupes)
#p = 0.8384
```

```{r warning = F, message = F}
keywords <- keywords %>%
  group_by(Keyword) %>% filter(n() >= 10)

countsKeywords <- keywords %>% 
  group_by(Keyword) %>%
  tally()

keywords %>%
  ggplot(aes(x = reorder(Keyword, `Open Rate`, FUN = median), y = `Open Rate`))+
  stat_boxplot(geom ='errorbar', width = .25) +
  geom_boxplot(fill = c('#003a5d', '#0091b3', '#003a5d', '#0091b3', '#003a5d', '#0091b3', '#003a5d', '#0091b3', '#003a5d', '#0091b3', '#003a5d'))+
  theme_ipsum(axis_title_size = 12)+
  theme(axis.text.x = element_text(angle = 90, hjust=1))+
  stat_summary(fun = median, geom = "crossbar", width = .75, color = "#a9ad00", size = .4)+
  labs(title = "Open Rate by Presence of Keywords",
       x = 'Keyword')+
  geom_hline(yintercept = median(emails$`Open Rate`), color = 'red', size = 1)+
  geom_text(data = countsKeywords,
            aes(Keyword, Inf, label = n), vjust = 1)

modelK <- lm(`Open Rate` ~ Keyword, keywords)
anovaK <- anova(modelK)
#p = 7.278912e-09

keywordTable <- keywords %>%
  group_by(Keyword)%>%
  summarize(
    `Open Rate` = median(`Open Rate`),
    CTOR = median(CTOR)
  )
```

```{r}
keywords %>%
  ggplot(aes(x = reorder(Keyword, CTOR, FUN = median), y = CTOR))+
  stat_boxplot(geom ='errorbar', width = .25) +
  geom_boxplot(fill = c('#003a5d', '#0091b3', '#003a5d', '#0091b3', '#003a5d', '#0091b3', '#003a5d', '#0091b3', '#003a5d', '#0091b3', '#003a5d'))+
  theme_ipsum(axis_title_size = 12)+
  theme(axis.text.x = element_text(angle = 90, hjust=1))+
  stat_summary(fun = median, geom = "crossbar", width = .75, color = "#a9ad00", size = .4)+
  labs(title = "Click-to-Open Rate by Presence of Keywords",
       x = 'Keyword')+
  geom_hline(yintercept = median(emails$`CTOR`), color = 'red', size = 1)+
  geom_text(data = countsKeywords,
            aes(Keyword, Inf, label = n), vjust = 1)
```

The keywords plot is interesting. For the sake of clutter, the white median text was removed, but a solid red line was added. This red line represents the median open or click-to-open rate for the respective plot. Some keywords like Help or Year-End stay in the same spot for open and click-to-open rate, while others move slightly. None, however, completely flip sides. In general, the same keywords are good for both open and click-to-open rate.

```{r warning = F, message = F}
# topics <- emails %>% 
#   group_by(Topic) %>% filter(n() >= 10)
# 
# countsTopics <- topics %>% 
#   group_by(Topic) %>%
#   tally()
# 
# topics %>%
#   ggplot(aes(x = reorder(Topic, `Open Rate`, FUN = median), y = `Open Rate`, fill = Topic))+
#   geom_boxplot(fill = c('#003a5d', '#0091b3', '#003a5d', '#0091b3', '#003a5d','#0091b3', '#003a5d', '#0091b3', '#003a5d', '#0091b3', '#003a5d', '#0091b3', '#003a5d', '#0091b3'))+
#   theme_ipsum(axis_title_size = 12)+
#   theme(axis.text.x = element_text(angle = 90, hjust=1),
#         legend.position = 'none')+
#   labs(title = 'Open Rate by Topic',
#        x = 'Topic')+
#   stat_summary(fun = median, geom = "crossbar", width = .75, color = "#a9ad00", size = .4)+
#   geom_hline(yintercept = median(emails$`Open Rate`), color = 'red', size = 1)
# 
# topicsTable <- topics %>%
#   group_by(Topic)%>%
#   summarize(
#     `Open Rate` = median(`Open Rate`),
#     CTOR = median(CTOR)
#   )
```

```{r}
# topics %>%
#   ggplot(aes(x = reorder(Topic, CTOR, FUN = median), y = CTOR, fill = Topic))+
#   geom_boxplot(fill = c('#003a5d', '#0091b3', '#003a5d', '#0091b3', '#003a5d','#0091b3', '#003a5d', '#0091b3', '#003a5d', '#0091b3', '#003a5d', '#0091b3', '#003a5d', '#0091b3'))+
#   theme_ipsum(axis_title_size = 12)+
#   theme(axis.text.x = element_text(angle = 90, hjust=1),
#         legend.position = 'none')+
#   labs(title = 'Click-to-Open Rate by Topic',
#        x = 'Topic')+
#   stat_summary(fun = median, geom = "crossbar", width = .75, color = "#a9ad00", size = .4)
```

```{r, warning =F, message =F}
#aes(x = reorder(`Product Type`, `Open Rate`, FUN = median), y = `Open Rate`, fill = `Product Type`)

# emails %>%
#   ggplot(aes(`Product Type`, `Open Rate`, fill = `Product Type`))+
#   geom_boxplot(fill = c('#003a5d', '#0091b3', '#003a5d', '#0091b3', '#003a5d'))+
#   theme_ipsum(axis_title_size = 12)+
#   theme(
#         legend.position = 'none')+
#   labs(title = 'Open Rate by Product Type',
#        x = 'Product Type')+
#   stat_summary(fun = median, geom = "crossbar", width = .75, color = "#a9ad00", size = .4)+
#   stat_summary(fun = median, geom = "text", aes(label = paste0(round(..y.., 1), "%")),
#                position = position_dodge(width = 0.75), vjust = -0.5, color = "white")

# emails$`Product Type` <- gsub("Visions", "Donor-Centric\nNewsletter", emails$`Product Type`)

# product_order <- c("Donor-Centric\nPostcard", "Donor-Centric\nTarget Mailer", "Donor-Centric\nNewsletter", "Professional\nAdvisor\nNewsletter")
# emails$`Product Type` <- factor(emails$`Product Type`, levels = product_order, ordered = TRUE)

# fs <- favstats(CTOR ~ `Product Type`, data = emails)
# 
# productTable <- emails %>%
#   group_by(`Product Type`) %>%
#   summarize(
#     `Open Rate` = median(`Open Rate`),
#     CTOR = median(CTOR)
#   )
```

```{r}
# emails %>%
#   ggplot(aes(`Product Type`, CTOR, fill = `Product Type`))+
#   geom_boxplot(fill = c('#003a5d', '#0091b3', '#003a5d', '#0091b3', '#003a5d'))+
#   theme_ipsum(axis_title_size = 12)+
#   theme(
#         legend.position = 'none')+
#   labs(title = 'Click-to-Open Rate by Product Type',
#        x = 'Product Type')+
#   stat_summary(fun = median, geom = "crossbar", width = .75, color = "#a9ad00", size = .4)
```

```{r warning = F, message = F}
# techs <- emails %>% 
#   group_by(Techniques) %>% 
#   summarise(
#     openRate=round(median(`Open Rate`), digits =4),
#     CTOR=round(median(CTOR), digits =4),
#     count=n()
#     )
# 
# techs <- techs[,c(-1, -4)]
# 
# chisqT <- chisq.test(techs)
# #p ≈ 0
```

```{r warning = F, message = F}
# techniques %>%
#   ggplot(aes(Day, `Open Rate`, fill = Day))+
#   geom_boxplot(fill = c('#003a5d', '#0091b3', '#003a5d', '#0091b3', '#003a5d'))+
#   theme_ipsum(axis_title_size = 12)+
#   labs(title = 'Open Rate by Day of the Week for Professional Advisor Newsletter Emails')+
#   theme(legend.position="none")+
#   stat_summary(fun = median, geom = "crossbar", width = .75, color = "#a9ad00", size = .4)+
#   stat_summary(fun = median, geom = "text", aes(label = paste0(round(..y.., 2), "%")), 
#                position = position_dodge(width = 0.75), vjust = -0.5, color = "white")
# 
# comparison1 <- emails %>% 
#   group_by(Day, Techniques) %>% 
#   summarise(
#     openRate=round(median(`Open Rate`), digits =4),
#     )
# 
# comparison1 <- comparison1 %>%
#   pivot_wider(id_cols = Techniques, names_from = Day, values_from = openRate)
# 
# comparison1 <- comparison1[-1,-2]
# 
# comparison2 <- emails %>% 
#   group_by(Day) %>% 
#   summarise(
#     openRate=round(median(`Open Rate`), digits =4),
#     )
# 
# comparison2 <- comparison2 %>%
#   pivot_wider(names_from = Day, values_from = openRate)
# 
# comparison1 <- comparison1[,-1]
# 
# comparison1 <- comparison1 %>%
#   mutate(source = 'Techniques')
# 
# comparison2 <- comparison2[,-1]
# 
# comparison2 <- comparison2 %>%
#   mutate(source = 'all')
# 
# comparison <- rbind(comparison1, comparison2)
# 
# comparison <- comparison[,-6]
# 
# chisqC <- chisq.test(comparison)
# #p = 0.9332
# 
# techniqueDayTable2 <- techniques %>%
#   group_by(Day) %>%
#   summarize(
#     `Open Rate` = median(`Open Rate`),
#     CTOR = median(CTOR)
#   )
```

```{r warning = F, message = F}
# countsTechniques2 <- techniques %>% 
#   group_by(Day, Covid) %>%
#   tally()
# 
# techniques %>%
#   ggplot(aes(Day, `Open Rate`, fill = Day,))+
#   geom_boxplot(fill = c('#003a5d', '#0091b3', '#003a5d', '#0091b3', '#003a5d', '#003a5d', '#0091b3', '#003a5d', '#0091b3', '#003a5d'))+
#   theme_ipsum(axis_title_size = 12) +
#     theme(
#       legend.position="none",
#       axis.text.x = element_text(angle = 90, hjust=1)
#     )+
#   labs(title = 'Open Rate by Day of the Week Relative to COVID-19 for Professional Advsior Newsletters')+
#   facet_wrap( ~ Covid)+
#   stat_summary(fun = median, geom = "crossbar", width = .75, color = "#a9ad00", size = .4)+
#   stat_summary(fun = median, geom = "text", aes(label = paste0(round(..y.., 1), "%")),
#                position = position_dodge(width = 0.75), vjust = -0.5, color = "white")
# 
# techPC <- emails %>% 
#   filter(`Product Type` == 'Professional\nAdvisor\nNewsletter' & Covid == 'Post-COVID')%>%
#   group_by(Day)%>%
#   summarise(
#     openRate = median(`Open Rate`),
#     CTOR = median(CTOR),
#     count=n()
#   )
# 
# techniqueDayTable <- techniques %>%
#   group_by(Covid, Day) %>%
#   summarize(
#     `Open Rate` = median(`Open Rate`),
#     CTOR = median(CTOR)
#   )
# 
# pre_covid <- techniqueDayTable[techniqueDayTable$Covid == "Pre-COVID", ]
# post_covid <- techniqueDayTable[techniqueDayTable$Covid == "Post-COVID", ]
# 
# # Merging on the day column
# merged_techniqueDayTable <- merge(pre_covid, post_covid, by = "Day", suffixes = c("_Pre", "_Post"))
# 
# # Calculating the differences
# merged_techniqueDayTable$click_rate_diff <- merged_techniqueDayTable$`CTOR_Post` - merged_techniqueDayTable$`CTOR_Pre`
# merged_techniqueDayTable$open_rate_diff <- merged_techniqueDayTable$`Open Rate_Post` - merged_techniqueDayTable$`Open Rate_Pre`
# 
# # Selecting the relevant columns
# final_techniqueDayTable <- merged_techniqueDayTable[, c("Day", "click_rate_diff", "open_rate_diff")]
```

```{r warning = F, message = F}
# orMax <- emails %>%
#   filter(Day == 'Friday' & Period =='Yes' & lastChar == 'Period' & Topic == 'Life Stages' & `Product Type` == 'Visions')
# 
# crMax <- emails %>%
#   filter(Day == 'Thursday' & Hyphen =='Yes' & lastChar == 'No Punctuation' & Topic == 'TCJA' & `Product Type` == 'Professional\nAdvisor\nNewsletter
# ')
```

```{r plots, warning = FALSE, message = F, include=F}
# both2 %>%
#   ggplot(aes(Source, `Open Rate`, fill = Source))+
#   geom_boxplot(fill = c('#003a5d', '#0091b3'))+
#   theme_ipsum(axis_title_size = 12) +
#     theme(
#       legend.position="none",
#     )+
#   labs(title = "Open Rate by Punctuation",
#        x = "Punctuation")+
#   stat_summary(fun = median, geom = "crossbar", width = .75, color = "#a9ad00", size = .4)+
#   stat_summary(fun = median, geom = "text", aes(label = paste0(round(..y.., 2), "%")), 
#                position = position_dodge(width = 0.75), vjust = -0.5, color = "white")
# 
# modelBoth2 <- lm(`Open Rate` ~ Source, both2)
# anovaB2 <- anova(modelBoth2)
# #p = 0.005394861
# 
# emails %>% 
#   ggplot(aes(Size, `Open Rate`, fill = Size))+
#   geom_boxplot(fill = c('#003a5d', '#0091b3'))+
#   theme_ipsum(axis_title_size = 12) +
#     theme(
#       legend.position="none",
#     )+
#   labs(title = "Open Rate by Charity Size")+
#   stat_summary(fun = median, geom = "crossbar", width = .75, color = "#a9ad00", size = .4)+
#   stat_summary(fun = median, geom = "text", aes(label = paste0(round(..y.., 2), "%")), 
#                position = position_dodge(width = 0.75), vjust = -0.5, color = "white")
# 
# sizeM <- lm(`Open Rate` ~ Size, emails)
# asizeM <- anova(sizeM)
# #p = 0.1473003
# 
# emails %>%
#   ggplot(aes(Duplicate, `Open Rate`))+
#   geom_boxplot(fill = c('#003a5d', '#0091b3'))+
#   theme_ipsum(axis_title_size = 12) +
#     theme(
#       legend.position="none",
#     )+
#   labs(title = "Open Rate by Duplicate Subjects")+
#   stat_summary(fun = median, geom = "crossbar", width = .75, color = "#a9ad00", size = .4)+
#   stat_summary(fun = median, geom = "text", aes(label = paste0(round(..y.., 2), "%")), 
#                position = position_dodge(width = 0.75), vjust = -0.5, color = "white")
# 
# modelD <- lm(`Open Rate` ~ Duplicate, data = emails)
# anovaD <- anova(modelD)
# #p = 0.7162413
```

```{r click plots, warning = F, message = F, include=F}
countsMF <- emailsMF %>% 
  group_by(Day) %>%
  tally()



counts7 <- havePunc %>% 
  group_by(Punc) %>%
  tally()



counts3 <- emails %>% 
  group_by(lastChar) %>%
  tally()

modelLC <- lm(CTOR ~ lastChar, emailsLC)
anovaLC <- anova(modelLC)

# countsTopics <- topics %>% 
#   group_by(Topic) %>%
#   tally()
# 
# techniques %>%
#   ggplot(aes(Day, CTOR, fill = Day))+
#   geom_boxplot(fill = c('#003a5d', '#0091b3', '#003a5d', '#0091b3', '#003a5d'))+
#   theme_ipsum(axis_title_size = 12)+
#   labs(title = 'Click-to-Open Rate by Day of the Week for Professional Advisor Newsletter Emails')+
#   theme(legend.position="none")+
#   stat_summary(fun = median, geom = "crossbar", width = .75, color = "#a9ad00", size = .4)+
#   stat_summary(fun = median, geom = "text", aes(label = paste0(round(..y.., 2), "%")), 
#                position = position_dodge(width = 0.75), vjust = -0.5, color = "white")
# 
# countsTechniques2 <- techniques %>% 
#   group_by(Day, Covid) %>%
#   tally()
# 
# techniques %>%
#   ggplot(aes(Day, CTOR, fill = Day,))+
#   geom_boxplot(fill = c('#003a5d', '#0091b3', '#003a5d', '#0091b3', '#003a5d', '#003a5d', '#0091b3', '#003a5d', '#0091b3', '#003a5d'))+
#   theme_ipsum(axis_title_size = 12) +
#     theme(
#       legend.position="none",
#       axis.text.x = element_text(angle = 90, hjust=1)
#     )+
#   labs(title = 'Click-to-Open Rate by Day of the Week Relative to COVID-19 for Professional Advsior Newsletters')+
#   facet_wrap( ~ Covid)+
#   stat_summary(fun = median, geom = "crossbar", width = .75, color = "#a9ad00", size = .4)+
#   stat_summary(fun = median, geom = "text", aes(label = paste0(round(..y.., 1), "%")),
#                position = position_dodge(width = 0.75), vjust = -0.5, color = "white")
```

# Conclusion

In order to maximize effectiveness, one should send on a Friday, with a subject line between 40-70 characters, use punctuation, but not dashes if looking for open rate, end with an exclamation point or a period, but preferably an exclamation point, and use keywords like Help, You/Your/Yours, and Impact/Effect. A table with all that information in a more concise way is below.

|      Maximize      |  Day   |  Subject Length  |    Punctuation?    |  Last Character   |     Keywords to Use      | Keywords to Avoid |
|:------------------:|:------:|:----------------:|:------------------:|:-----------------:|:------------------------:|:-----------------:|
|     Open Rate      | Friday | 40-70 characters | Yes, except dashes | Exclamation Point | Help, You, Impact/Effect |  Year-End, Easy   |
| Click-to-Open Rate | Friday | 40-70 characters |        Yes         | Exclamation Point |      Help, You, Tax      |  Year-End, Easy   |
